Application of Machine Learning to Predict the Optoelectronic Properties of Benzobisazoles

### 2014-08-15 paper/paper.tex
More paper updates

\begin{abstract}
\hl{Because there are always more molecules to test than time}, there is a need to quickly be able to rule out structures that are unsuitable for further research. The method presented in this paper takes advantage of common structural features of molecules to decrease the required computational expense that can be seen equivalent Density Functional Theory calculations. By using machine learning methods, the HOMO, LUMO, and band gap can be predicted to within 0.09 eV, 0.08 eV, and 0.10 eV respectively for the benzobisazole structures tested. These results also come with the added benefit of being 30,000,000 times faster than the equivalent DFT calculations.
\end{abstract}

@@

To take on both of both of these problems, we will be looking machine learning methods to both be able to predict molecular properties, but with the goal that these calculations would be inexpensive but retain the same level of accuracy. In recent work, various machine learning methods have been shown to be fast ways to predict properties of many systems and specifically, has been shown to be a viable way to predict properties of molecules\cite{hansen_assessment_2013} \cite{montavon_machine_2013}.

@@

@@

In the work by Montavon et al., it has been shown that different machine learning algorithms have had drastically different performances when used to predict molecular properties. Because of this, seven algorithms were picked for comparison. These were mean value, linear regression, linear ridged regression, tree regression, k-nearest neighbors, and support vector machines. The support vector machines method was then broken down into two methods, one using a Gaussian kernel, and the other using a Laplacian kernel. The mean value method is used as a baseline for all of the other methods with its interpretation being the error that would be gotten by just guessing.

@@

 \begin{figure}[H]
   \begin{center}
     \includegraphics[width=5in]{figures/surf.png}
   \end{center}
   \caption{Example scan of test errors (eV) as a function of SVM hyperparameters C and $\gamma$. The minimum located in the center corresponding to the ideal pair of hyperparameter values to use for the machine learning model. Similar scan procedures were done for the hyperparameters of the other models to find the lowest testing error.}
   \label{fig:scan}
 \end{figure}

@@

@@

Once all the data and feature vectors were collected, they were feed into the machine learning algorithms. For each feature vector, property, method set the methods relevant hyperparameters were optimized (C, $\gamma$, max depth, $\alpha$, n neighbors). Then using these optimized hyperparameters, the machine learning models were trained using k-folds for cross validation, specifically using 10 folds \hl{[CITE]}. All of the results were then tabulated and compared based on their mean absolute errors (MAE) calculated as the distance the ML predicted value was from DFT calculated value.

@@

For the feature vectors, the decay was the best performing feature vector with the parameters $A = 1, H = 1, p = 1$. \hl{(0.71)} As well as having a better overall performance, it is also much faster than the naive feature vector because it has fewer components. This increased speed could especially be seen when training the data. This can be attributed to the time complexity of $O(n^2) or O(n^3)$ where n is the number of dimensions.

@@

\hl{David Yaron, Gregorie Montavon, Bucknell (Marcy), University of North Georgia Department of Chemistry and Biochemistry, ...}

@@

>>>>

\begin{abstract}
\hl{Because there are always more molecules to test than time}, there is a need to quickly be able to rule out structures that are unsuitable for further research. The method presented in this paper takes advantage of common structural features of molecules to decrease the required computational expense that can be seen equivalent Density Functional Theory calculations. By using supervised machine learning methods, the HOMO, LUMO, and band gap can be predicted to within 0.09 eV, 0.08 eV, and 0.10 eV respectively for the benzobisazole structures tested. These results also come with the added benefit of being 30,000,000 times faster than the equivalent DFT calculations.
\end{abstract}

@@

To take on both of both of these problems, we will be looking machine learning methods to both be able to predict molecular properties inexpensively but retain the same level of accuracy as DFT. In recent work, various machine learning methods have been shown to be fast ways to predict properties of many systems and specifically, has been shown to be a viable way to predict properties of molecules\cite{hansen_assessment_2013} \cite{montavon_machine_2013}.

@@

\subsection{Coulomb Matrix}
(Montavon et al.)
$$
M_{ij} =
\begin{cases}
    0.5 Z_i^{2.4} &: i = j \\
    \frac{Z_i Z_j}{|R_i - R_j|} &: i \ne j \\
\end{cases}
$$
Specifically, to reduce the dimensionality Principal Component Analysis was used to reduce these vectors from 52,000 elements to 500 elements.

(Take lower triangle and convert the matrix to a vector)
$$M \rightarrow M_v$$
$$PCA(M_v) \rightarrow M_{smaller}$$

\hl{Should I add this and include those results?}

@@

In the work by Montavon et al., it has been shown that different machine learning algorithms have had drastically different performances when used to predict molecular properties. Because of this, seven algorithms were picked for comparison. These were mean value, linear regression, linear ridged regression, tree regression, k-nearest neighbors, and support vector machines. The support vector machines method was then broken down into two methods, one using a Gaussian kernel, and the other using a Laplacian kernel. The mean value method is used as a baseline for all of the other methods with its interpretation being the error that would be gotten by just guessing. The mean value method does not use at the feature vector to make its prediction, so all of the different methods will have the same result for it. \hl{Should I add NN?}

@@

 \begin{figure}[H]
   \begin{center}
     \includegraphics[width=5in]{figures/surf.png}
   \end{center}
   \caption{Example scan of test errors (eV) as a function of SVM hyperparameters C and $\gamma$. The minimum located in the center corresponding to the ideal pair of hyperparameter values to use for the machine learning model. Similar scan procedures were done for the hyperparameters of the other models to find the lowest testing error.\hl{Relevant?}}
   \label{fig:scan}
 \end{figure}

@@

 \begin{figure}[H]
   \begin{center}
     \includegraphics[width=5in]{figures/overall.pdf}
   \end{center}
   \caption{\hl{include something similar?}}
   \label{fig:scan}
 \end{figure}

@@

Once all the data and feature vectors were collected, they were feed into the machine learning algorithms. Specifically, we will be looking at supervised learning, meaning that the methods will be trained on (feature vector, property) pairs and then for the testing they will be given just feature vectors and be expected to estimate the respective property value. For each feature vector, property, method set the methods relevant hyperparameters were optimized (C, $\gamma$, max depth, $\alpha$, n neighbors). Then using these optimized hyperparameters, the machine learning models were trained using k-folds for cross validation, specifically using 10 folds \hl{[CITE]}. k-Folds and the cross validation are used to ensure that the final model generated is not biased towards the particular set of data used for training (overfitting). All of the results were then tabulated and compared based on their mean absolute errors (MAE) calculated as the distance the ML predicted value was from DFT calculated value.

@@

For the feature vectors, the decay was the best performing feature vector with the parameters $A = 1, H = 1, p = 1$. \hl{(0.71)} As well as having a better overall performance, it is also much faster than the naive feature vector because it has fewer components. This increased speed could especially be seen when training the data. \hl{This can be attributed to the time complexity of $O(n^2) or O(n^3)$ where n is the number of dimensions.}

@@

\hl{David Yaron, Gregorie Montavon, Bucknell (Marcy), XSEDE (Blacklight, Gordon, Trestles), University of North Georgia Department of Chemistry and Biochemistry, ...}

@@

All of the code and data for this paper can be found in the following two git repositories.

All of the actual machine learning code:

\url{http://www.github.com/crcollins/ML/}

The code to generate benzobisazole structures/names:

\url{http://www.github.com/crcollins/chemtools-webapp/}
